---
title: "Statistical Calculation Lab1"
author: "ID NAME"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1. Install R and Rstudio
# 2. Random Number Generation and Analysis Report
## 2.0. Set Seed and Preparation Work
```{r prepare, message=FALSE, warning=FALSE}
set.seed(1011)
library(rticles)
library(ggplot2)
library(gridExtra)
library(dplyr)
```
## 2.1. Generate 1000 t-distribution random numbers with 1 degree of freedom
```{r, message=FALSE, warning=FALSE}
t_random <- rt(1000, df = 1)
# Create scatter plot and histogram
p1 <- ggplot(data.frame(x = 1:1000, y = t_random), aes(x = x, y = y)) +
  geom_point(alpha = 0.6, color = "blue") +
  labs(title = "t(1) Distribution Random Numbers Scatter Plot",
       x = "Index",
       y = "Random Value") +
  theme_minimal()
p2 <- ggplot(data.frame(x = t_random), aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightblue",
                 color = "black",
                 alpha = 0.7) +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "t(1) Distribution Random Numbers Histogram",
       x = "Random Value",
       y = "Density") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 1)
```
## 2.2. Generate 1000 F-distribution random numbers with degrees of freedom 2 and 5
```{r, message=FALSE, warning=FALSE}
f_random <- rf(1000, df1 = 2, df2 = 5)
# Create scatter plot and histogram
p3 <- ggplot(data.frame(x = 1:1000, y = f_random), aes(x = x, y = y)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  labs(title = "F(2,5) Distribution Random Numbers Scatter Plot",
       x = "Index",
       y = "Random Value") +
  theme_minimal()
p4 <- ggplot(data.frame(x = f_random), aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightgreen",
                 color = "black",
                 alpha = 0.7) +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "F(2,5) Distribution Random Numbers Histogram",
       x = "Random Value",
       y = "Density") +
  theme_minimal() +
  xlim(0, quantile(f_random, 0.99))  # Limit x-axis range to better display main distribution
grid.arrange(p3, p4, ncol = 1)
```
## 2.3. Mixed Normal Distribution Random Numbers
```{r, message=FALSE, warning=FALSE}
# Generate 200 random numbers: 100 N(0,1) and 100 N(1,5)
normal_random1 <- rnorm(100, mean = 0, sd = 1)    # N(0,1)
normal_random2 <- rnorm(100, mean = 1, sd = sqrt(5))  # N(1,5), note: sd is standard deviation
# Combine two sets of random numbers
mixed_normal <- c(normal_random1, normal_random2)
# Create group labels
groups <- factor(c(rep("N(0,1)", 100), rep("N(1,5)", 100)))
# Create data frame
normal_df <- data.frame(
  values = mixed_normal,
  group = groups,
  index = 1:200
)
# Create scatter plot (colored by group)
p5 <- ggplot(normal_df, aes(x = index, y = values, color = group)) +
  geom_point(alpha = 0.7) +
  labs(title = "Mixed Normal Distribution Random Numbers Scatter Plot",
       x = "Index",
       y = "Random Value",
       color = "Distribution Type") +
  scale_color_manual(values = c("red", "blue")) +
  theme_minimal()
# Create histogram
p6 <- ggplot(data.frame(values = mixed_normal), aes(x = values)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "purple",
                 color = "black",
                 alpha = 0.7) +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "Mixed Normal Distribution - Combined Histogram",
       x = "Random Value",
       y = "Density") +
  theme_minimal()
grid.arrange(p5, p6, ncol = 1)
```
# 3. Exercises of Chapter 1
## 3.1 Exercise 1: Pareto Distribution
```{r, message=FALSE, warning=FALSE}
a <- 2
b <- 2
n <- 1000
u <- runif(n)
pareto_random <- b / (1 - u)^(1/a)
pareto_df <- data.frame(x = pareto_random)
pareto_density <- function(x) {
  ifelse(x >= b, (a * b^a) / (x^(a + 1)), 0)
}
ggplot(pareto_df, aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = pareto_density,
                color = "red",
                linewidth = 1.2,
                aes(color = "Theoretical Density")) +
  labs(title = "Pareto(2,2) Distribution",
       x = "Value",
       y = "Density") +
  scale_color_manual(name = "", values = c("Theoretical Density" = "red")) +
  theme_minimal() +
  theme(legend.position = "top") +
  xlim(b, quantile(pareto_random, 0.99))
```
## 3.2 Exercise 2: Mixed Normal Distribution
```{r, message=FALSE, warning=FALSE}
n<-1000
eta<-NULL
i<-1
p<-0.5
while(i<=n){
  r<-runif(1,0,1)
  if(r<p){
    eta[i]<-rnorm(1,0,1)
  }
  else{
    eta[i]<-rnorm(1,3,1)
  }
  i=i+1
}
mixed_normal_density <- function(x) {
  comp1 <- dnorm(x, mean = 0, sd = 1)
  comp2 <- dnorm(x, mean = 3, sd = 1)
  0.5 * comp1 + 0.5 * comp2
}
ggplot(data.frame(x = eta), aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = mixed_normal_density,
                color = "red",
                linewidth = 1.2,
                aes(color = "Theoretical Density")) +
  labs(title = "Mixed Normal Distribution",
       x = "Value",
       y = "Density") +
  scale_color_manual(name = "", values = c("Theoretical Density" = "red")) +
  theme_minimal() +
  theme(legend.position = "top") +
  xlim(min(eta), max(eta))
```
## 3.3 Exercise 3: Rayleigh Distribution
(1) Inverse transformation sampling:
```{r, message=FALSE, warning=FALSE}
sigma <- 1
n <- 1000
u <- runif(n)
rayleigh_random <- sigma*sqrt(-2*log(1-u))
rayleigh_df <- data.frame(x = rayleigh_random)
rayleigh_density <- function(x) {
  ifelse(x >= 0, x/(sigma^2)*exp(-x^2/(2*sigma^2)), 0)
}
ggplot(rayleigh_df, aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = rayleigh_density,
                color = "red",
                linewidth = 1.2,
                aes(color = "Theoretical Density")) +
  labs(title = "Rayleigh Distribution - Inverse Transformation Sampling",
       x = "Value",
       y = "Density") +
  scale_color_manual(name = "", values = c("Theoretical Density" = "red")) +
  theme_minimal() +
  theme(legend.position = "top") +
  xlim(0, quantile(rayleigh_random, 0.99))
```
(2) Transformed sampling
```{r, message=FALSE, warning=FALSE}
sigma <- 1
n <- 1000
i<-1
eta = NULL
while(i<=n){
  x1<-rnorm(1,0,1)
  x2<-rnorm(1,0,1)
  eta[i]<-sqrt(x1^2+x2^2)
  i=i+1
}
rayleigh_df <- data.frame(x = eta)
rayleigh_density <- function(x) {
  ifelse(x >= 0, x/(sigma^2)*exp(-x^2/(2*sigma^2)), 0)
}
ggplot(rayleigh_df, aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "lightblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = rayleigh_density,
                color = "red",
                linewidth = 1.2,
                aes(color = "Theoretical Density")) +
  labs(title = "Rayleigh Distribution - Transformed Sampling",
       x = "Value",
       y = "Density") +
  scale_color_manual(name = "", values = c("Theoretical Density" = "red")) +
  theme_minimal() +
  theme(legend.position = "top") +
  xlim(0, quantile(rayleigh_random, 0.99))
```
## 3.4 Exercise 4: Distribution of Discrete Random Variable
```{r, message=FALSE, warning=FALSE}
n<-1000
i<-1
x<-0:4
p<-c(0.1,0.2,0.2,0.2,0.3)
eta_values <- numeric(n)
while(i <= n){
  sum_val <- 0
  r <- runif(1)
  for(j in 1:6){
    if(r < sum_val){
      break
    }
    else{
      sum_val <- sum_val + p[j]
    }
  }
  eta_values[i] <- x[j-1]
  i <- i + 1
}
observed_freq <- table(eta_values) / n
plot_data <- data.frame(
  Value = x,
  Theoretical = p,
  Observed = as.numeric(observed_freq[as.character(x)])
)
ggplot(plot_data) +
  geom_bar(aes(x = factor(Value), y = Observed, fill = "Observed Frequency"),
           stat = "identity", alpha = 0.7) +
  geom_line(aes(x = factor(Value), y = Theoretical, group = 1, color = "Theoretical Probability"),
            size = 1.5) +
  geom_point(aes(x = factor(Value), y = Theoretical, color = "Theoretical Probability"),
             size = 3) +
  scale_fill_manual(name = "", values = c("Observed Frequency" = "lightblue")) +
  scale_color_manual(name = "", values = c("Theoretical Probability" = "red")) +
  labs(title = "Distribution of Discrete Random Variable",
       x = "Value",
       y = "Frequancy/Probability") +
  theme_minimal() +
  theme(legend.position = "top")
```
# 4. Classroom Practices
## Practice 1:
```{r, message=FALSE, warning=FALSE}
# practice 1:
vec <- seq(from = 10, to = 30, by = 3)
print(vec)
```
## Practice 2:
```{r, message=FALSE, warning=FALSE}
mat <- matrix(x, nrow = 2, ncol = 3, byrow = TRUE)
rownames(mat) <- c("row1", "row2")
colnames(mat) <- c("C.1", "C.2", "C.3")
print(mat)
```