---
title: "Statistical Calculation Lab4"
author: "ID NAME"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rticles)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(reshape2)
library(tidyr)
```
# 1. Recurrence of Textbook Examples
## Example 6.1.1: Bootstrap estimation of the bias of variance estimators
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n=100; mu=0; sigma=2; m=2000;
X=rnorm(n,mu,sigma);
sigma2hat=mean((X-mean(X))^2)
sigma2Bst=NULL
for (i in (1:m)){
  O=sample(seq(1,n),n,replace=T)
  Bstrap_X=X[O]
  sigma2Bst[i]=mean((Bstrap_X-mean(Bstrap_X))^2)
}
Bias=mean(sigma2Bst)-sigma2hat
cat(Bias)
```
## Example 6.1.3: Bootstrap estimation of the bias of the mean difference estimator
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n1=100
n2=200
mu1=0
mu2=0.3
sigma1=1
sigma2=2
X=rnorm(n1,mu1,sigma1)
Y=rnorm(n2,mu2,sigma2)
diffhat=mean(X)-mean(Y)
m=2000
diffBst=NULL
for (i in (1:m)){
  O1=sample(seq(1,n1),n1,replace=T)
  O2=sample(seq(1,n2),n2,replace=T)
  Bst_X=X[O1]
  Bst_Y=Y[O2]
  diffBst[i]=mean(Bst_X)-mean(Bst_Y)
}
Bias=mean(diffBst)-diffhat
cat(Bias)
```
## Example 6.1.5: Bootstrap estimation of the variance of the correlation coefficient estimator
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
library(mnormt)
n=100
mu1=mu2=0
sigma1=1
sigma2=2
rho=0.6
sigma12=sigma21=rho*sigma1*sigma2
mu=c(mu1,mu2)
Sigma=matrix(c(sigma1^2,sigma12,sigma21,sigma2^2),2,2)
Data=rmnorm(n,mu,Sigma)
X=Data[,1]
Y=Data[,2]
rhohat=sum((X-mean(X))*(Y-mean(Y)))/sqrt(sum((X-mean(X))^2)*sum((Y-mean(Y))^2));
m=2000
rhoBst=NULL
for (i in (1:m)){
  O=sample(seq(1,n),n,replace=T);
  Bst_Data=Data[O,]
  X=Bst_Data[,1]
  Y=Bst_Data[,2]
  rhoBst[i]=sum((X-mean(X))*(Y-mean(Y)))/sqrt(sum((X-mean(X))^2)*sum((Y-mean(Y))^2))
}
var_rhoBst=mean((rhoBst-mean(rhoBst))^2)
cat(var_rhoBst)
```
## Example 6.2.1: Jackknife estimation of the bias of variance estimators
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n=100
mu=0
sigma=2
X=rnorm(n,mu,sigma)
sigma2hat=sum((X-mean(X))^2)/n
sigma2Jack=NULL
for (i in (1:n)){
  Jack_X=X[-i]
  sigma2Jack[i]=sum((Jack_X-mean(Jack_X))^2)/(n-1)
}
Bias=(n-1)*mean(sigma2Jack-sigma2hat)
cat(Bias)
```
## Example 6.2.4: Jackknife estimation of the correlation coefficient and its variance
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
library(mnormt)
n=100
mu1=mu2=0
sigma1=1
sigma2=2
rho=0.6
sigma12=sigma21=rho*sigma1*sigma2
mu=c(mu1,mu2)
Sigma=matrix(c(sigma1^2,sigma12,sigma21,sigma2^2),2,2)
Data=rmnorm(n,mu,Sigma)
X=Data[,1]
Y=Data[,2]
rhohat=sum((X-mean(X))*(Y-mean(Y)))/ sqrt(sum((X-mean(X))^2)*sum((Y-mean(Y))^2))
rhoJack=NULL
for (i in (1:n)){
  Jack_Data=Data[-i,]
  X=Jack_Data[,1]
  Y=Jack_Data[,2]
  rhoJack[i]=sum((X-mean(X))*(Y-mean(Y)))/sqrt(sum((X-mean(X))^2)*sum((Y-mean(Y))^2))
}
Varhat=(n-1)*mean((rhoJack-mean(rhoJack))^2)
cat(Varhat)
```
## Example 6.3.1: Bootstrap estimation of confidence intervals
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
alpha=0.05
n=100
mu=0
sigma=1
X=rnorm(n,mu,sigma)
muhat=mean(X)
m=K=rep_muBst=500
muBst=t=NULL
for (i in (1:m)){
  O=sample(seq(1,n),n,replace=T)
  Bstrap_X=X[O]
  muBst[i]=mean(Bstrap_X)
  for (k in (1:K)){
    O=sample(seq(1,n),n,replace=T)
    rep_Bstrap_X=Bstrap_X[O]
    rep_muBst[k]=mean(rep_Bstrap_X)
  }
  se_rep=sqrt(mean((rep_muBst-mean(rep_muBst))^2))
  t[i]=(muBst[i]-muhat)/se_rep
}
se_muBst=sqrt(mean((muBst-mean(muBst))^2))
L1=muhat-qnorm(1-alpha/2,0,1)*se_muBst
U1=muhat+qnorm(1-alpha/2,0,1)*se_muBst
smuBst=sort(muBst)
L2=smuBst[round(m*alpha/2)]
U2=smuBst[round(m*(1-alpha/2))]
L3=2*muhat-smuBst[round(m*(1-alpha/2))]
U3=2*muhat-smuBst[round(m*alpha/2)]
st=sort(t)
L4=muhat-st[round(m*(1-alpha/2))]*se_muBst
U4=muhat-st[round(m*alpha/2)]*se_muBst
c(L1,U1)
c(L2,U2)
c(L3,U3)
c(L4,U4)
```
# 2. Exercises
## Exercise 6.1: Bootstrap method
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n <- 100
m <- 200
u1 <- 0
u2 <- 1
sigma1 <- 1
sigma2 <- 2
B <- 1000  # Number of bootstrap replicates
# Generate original samples
x_original <- rnorm(n, u1, sigma1)
y_original <- rnorm(m, u2, sigma2)
# True parameter
true_ratio <- (sigma1^2) / (sigma2^2)
# Function to calculate variance ratio
calc_ratio <- function(x, y) {
  var(x) / var(y)
}
# Original estimate
original_estimate <- calc_ratio(x_original, y_original)
# Bootstrap procedure
bootstrap_estimates <- numeric(B)
for (i in 1:B) {
  x_boot <- sample(x_original, n, replace = TRUE)
  y_boot <- sample(y_original, m, replace = TRUE)
  bootstrap_estimates[i] <- calc_ratio(x_boot, y_boot)
}
# Calculate bias and variance
bootstrap_bias <- mean(bootstrap_estimates) - original_estimate
bootstrap_variance <- var(bootstrap_estimates)
cat("True ratio:", true_ratio, "\n")
cat("Original estimate:", original_estimate, "\n")
cat("Bootstrap bias:", bootstrap_bias, "\n")
cat("Bootstrap variance:", bootstrap_variance, "\n")
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df_boot <- data.frame(estimates = bootstrap_estimates)
ggplot(df_boot, aes(x = estimates)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", alpha = 0.7) +
  geom_density(color = "darkblue", linewidth = 1) +
  geom_vline(xintercept = original_estimate, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = true_ratio, color = "green", linetype = "dashed", linewidth = 1) +
  labs(title = "Bootstrap Distribution of Variance Ratio Estimate",
       x = "Variance Ratio Estimates",
       y = "Density",
       subtitle = paste("Bias =", round(bootstrap_bias, 4),
                       "Variance =", round(bootstrap_variance, 4))) +
  theme_minimal()
```
## Exercise 6.2: Jackknife method
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n <- 100
m <- 200
u1 <- 0
u2 <- 1
sigma1 <- 1
sigma2 <- 2
# Generate original samples
x_original <- rnorm(n, u1, sigma1)
y_original <- rnorm(m, u2, sigma2)
# Function to calculate variance ratio
calc_ratio <- function(x, y) {
  var(x) / var(y)
}
# Original estimate
original_estimate <- calc_ratio(x_original, y_original)
# Jackknife procedure - delete X observations
jackknife_x <- numeric(n)
for (i in 1:n) {
  x_jack <- x_original[-i]
  jackknife_x[i] <- calc_ratio(x_jack, y_original)
}
# Jackknife procedure - delete Y observations
jackknife_y <- numeric(m)
for (j in 1:m) {
  y_jack <- y_original[-j]
  jackknife_y[j] <- calc_ratio(x_original, y_jack)
}
# Combine jackknife estimates
all_jackknife <- c(jackknife_x, jackknife_y)
pseudo_values <- (n + m) * original_estimate - (n + m - 1) * all_jackknife
# Calculate jackknife estimates
jackknife_estimate <- mean(pseudo_values)
jackknife_bias <- (n + m - 1) * (mean(all_jackknife) - original_estimate)
jackknife_variance <- var(pseudo_values) / (n + m)
cat("True ratio:", (sigma1^2) / (sigma2^2), "\n")
cat("Original estimate:", original_estimate, "\n")
cat("Jackknife bias:", jackknife_bias, "\n")
cat("Jackknife variance:", jackknife_variance, "\n")
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df_jack <- data.frame(
  type = rep(c("Delete X", "Delete Y"), c(n, m)),
  estimates = c(jackknife_x, jackknife_y)
)
ggplot(df_jack, aes(x = estimates, fill = type)) +
  geom_histogram(aes(y = ..density..), alpha = 0.6, position = "identity", bins = 30) +
  geom_vline(xintercept = original_estimate, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = (sigma1^2) / (sigma2^2), color = "green", linetype = "dashed", linewidth = 1) +
  labs(title = "Jackknife Distribution of Variance Ratio Estimates",
       x = "Variance Ratio Estimates",
       y = "Density",
       fill = "Jackknife Type",
       subtitle = paste("Bias =", round(jackknife_bias, 4),
                       "Variance =", round(jackknife_variance, 4))) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
## Exercise 6.3: Bootstrap method estimates the parameters of the Gamma distribution
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n <- 200
a_true <- 2
lambda_true <- 1
B <- 1000
# Generate original sample from Gamma distribution
x_original <- rgamma(n, shape = a_true, rate = lambda_true)
# Method of moments estimators for Gamma distribution
gamma_estimators <- function(x) {
  mean_x <- mean(x)
  var_x <- var(x)
  a_hat <- mean_x^2 / var_x
  lambda_hat <- mean_x / var_x
  return(c(a_hat, lambda_hat))
}
# Original estimates
original_estimates <- gamma_estimators(x_original)
# Bootstrap procedure
bootstrap_estimates <- matrix(0, nrow = B, ncol = 2)
for (i in 1:B) {
  x_boot <- sample(x_original, n, replace = TRUE)
  bootstrap_estimates[i, ] <- gamma_estimators(x_boot)
}
# Calculate bias and variance
bootstrap_bias <- colMeans(bootstrap_estimates) - original_estimates
bootstrap_variance <- apply(bootstrap_estimates, 2, var)
cat("True parameters: a =", a_true, "lambda =", lambda_true, "\n")
cat("Original estimates: a_hat =", original_estimates[1], "lambda_hat =", original_estimates[2], "\n")
cat("Bootstrap bias - a:", bootstrap_bias[1], "lambda:", bootstrap_bias[2], "\n")
cat("Bootstrap variance - a:", bootstrap_variance[1], "lambda:", bootstrap_variance[2], "\n")
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df_gamma <- data.frame(
  a_estimates = bootstrap_estimates[, 1],
  lambda_estimates = bootstrap_estimates[, 2]
)
# Plot for shape parameter a
p1 <- ggplot(df_gamma, aes(x = a_estimates)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightcoral", alpha = 0.7) +
  geom_density(color = "darkred", linewidth = 1) +
  geom_vline(xintercept = a_true, color = "green", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = original_estimates[1], color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Bootstrap Distribution of Shape Parameter (a)",
       x = "Shape Parameter Estimates",
       y = "Density") +
  theme_minimal()
# Plot for rate parameter lambda
p2 <- ggplot(df_gamma, aes(x = lambda_estimates)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightgreen", alpha = 0.7) +
  geom_density(color = "darkgreen", linewidth = 1) +
  geom_vline(xintercept = lambda_true, color = "green", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = original_estimates[2], color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Bootstrap Distribution of Rate Parameter (位)",
       x = "Rate Parameter Estimates",
       y = "Density") +
  theme_minimal()
# Display both plots
library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```
## Exercise 6.5: Jackknife after Bootstrap
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n <- 200
sigma_true <- 2
B <- 1000
# Generate original sample
x_original <- rnorm(n, 0, sigma_true)
# Function for variance estimator
variance_estimator <- function(x) {
  mean((x - mean(x))^2)
}
# Original estimate
original_variance <- variance_estimator(x_original)
# Bootstrap procedure for bias estimation
bootstrap_variances <- numeric(B)
for (i in 1:B) {
  x_boot <- sample(x_original, n, replace = TRUE)
  bootstrap_variances[i] <- variance_estimator(x_boot)
}
# Bootstrap bias estimate
bootstrap_bias_estimate <- mean(bootstrap_variances) - original_variance
# Jackknife procedure to estimate variance of bootstrap bias estimate
jackknife_bias_estimates <- numeric(n)
for (i in 1:n) {
  x_jack <- x_original[-i]
  # Perform bootstrap on jackknife sample
  jack_bootstrap_vars <- numeric(100)  # Reduced for computational efficiency
  for (j in 1:100) {
    x_jack_boot <- sample(x_jack, n-1, replace = TRUE)
    jack_bootstrap_vars[j] <- variance_estimator(x_jack_boot)
  }
  jack_original_var <- variance_estimator(x_jack)
  jackknife_bias_estimates[i] <- mean(jack_bootstrap_vars) - jack_original_var
}
# Jackknife variance estimate for bootstrap bias
jackknife_variance_bias <- ((n - 1) / n) * sum((jackknife_bias_estimates - mean(jackknife_bias_estimates))^2)
cat("True variance:", sigma_true^2, "\n")
cat("Original variance estimate:", original_variance, "\n")
cat("Bootstrap bias estimate:", bootstrap_bias_estimate, "\n")
cat("Jackknife variance of bootstrap bias estimate:", jackknife_variance_bias, "\n")
cat("Theoretical bias:", -sigma_true^2/n, "\n")
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df_comparison <- data.frame(
  Bootstrap_Variances = bootstrap_variances,
  Jackknife_Bias_Estimates = jackknife_bias_estimates
)
# Plot bootstrap distribution
p_boot <- ggplot(df_comparison, aes(x = Bootstrap_Variances)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", alpha = 0.7) +
  geom_density(color = "darkblue", linewidth = 1) +
  geom_vline(xintercept = sigma_true^2, color = "green", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = original_variance, color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Bootstrap Distribution of Variance Estimates",
       x = "Variance Estimates",
       y = "Density") +
  theme_minimal()
# Plot jackknife distribution of bias estimates
p_jack <- ggplot(data.frame(Bias_Estimates = jackknife_bias_estimates),
                 aes(x = Bias_Estimates)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightyellow", alpha = 0.7) +
  geom_density(color = "darkorange", linewidth = 1) +
  geom_vline(xintercept = bootstrap_bias_estimate, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = -sigma_true^2/n, color = "green", linetype = "dashed", linewidth = 1) +
  labs(title = "Jackknife Distribution of Bootstrap Bias Estimates",
       x = "Bias Estimates",
       y = "Density",
       subtitle = paste("Variance of bias estimates =", round(jackknife_variance_bias, 6))) +
  theme_minimal()
grid.arrange(p_boot, p_jack, ncol = 1)
```
# 3. Classroom Practices:
## Practice 1: Confidence interval of the sample median
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
sample_data <- rnorm(50, mean = 10, sd = 2)
n_bootstrap = 1000
conf_level = 0.95
n <- length(sample_data)
bootstrap_medians <- numeric(n_bootstrap)
for(i in 1:n_bootstrap) {
  bootstrap_sample <- sample(sample_data, n, replace = TRUE)
  bootstrap_medians[i] <- median(bootstrap_sample)
}
alpha <- 1 - conf_level
ci_lower <- quantile(bootstrap_medians, alpha/2)
ci_upper <- quantile(bootstrap_medians, 1 - alpha/2)
cat(ci_lower, ci_upper)
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df <- data.frame(medians = bootstrap_medians)
ggplot(df, aes(x = medians)) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", alpha = 0.7) +
    geom_density(color = "blue") +
    geom_vline(xintercept = ci_lower, color = "red", linetype = "dashed") +
    geom_vline(xintercept = ci_upper, color = "red", linetype = "dashed") +
    geom_vline(xintercept = median(sample_data), color = "darkgreen", size = 1) +
    labs(title = "Bootstrap Distribution of Median",
         x = "Bootstrap Medians", y = "Density") +
    theme_minimal()
```
## Practice 2: Correlation coefficient
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
x <- c(2, 3, 5, 7, 9, 11, 12, 14, 15, 16, 20)
y <- c(55, 62, 68, 75, 79, 85, 82, 88, 90, 92, 65)
n_bootstrap = 1000
conf_level = 0.95
n <- length(x)
pearson_boot <- numeric(n_bootstrap)
spearman_boot <- numeric(n_bootstrap)
for(i in 1:n_bootstrap) {
  indices <- sample(1:n, n, replace = TRUE)
  x_boot <- x[indices]
  y_boot <- y[indices]
  pearson_boot[i] <- cor(x_boot, y_boot, method = "pearson")
  spearman_boot[i] <- cor(x_boot, y_boot, method = "spearman")
}
pearson_original <- cor(x, y, method = "pearson")
spearman_original <- cor(x, y, method = "spearman")
pearson_bias <- mean(pearson_boot) - pearson_original
spearman_bias <- mean(spearman_boot) - spearman_original
pearson_se <- sd(pearson_boot)
spearman_se <- sd(spearman_boot)
alpha <- 1 - conf_level
pearson_ci <- quantile(pearson_boot, c(alpha/2, 1 - alpha/2))
spearman_ci <- quantile(spearman_boot, c(alpha/2, 1 - alpha/2))
results_df <- data.frame(
  Method = c("Pearson", "Spearman"),
  Original = c(pearson_original, spearman_original),
  Bias = c(pearson_bias, spearman_bias),
  Standard_Error = c(pearson_se, spearman_se),
  CI_Lower = c(pearson_ci[1], spearman_ci[1]),
  CI_Upper = c(pearson_ci[2], spearman_ci[2])
)
print(results_df)
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df <- data.frame(
    Correlation = c(pearson_boot, spearman_boot),
    Method = rep(c("Pearson", "Spearman"), each = n_bootstrap)
  )
ggplot(df, aes(x = Correlation, fill = Method)) +
    geom_density(alpha = 0.5) +
    geom_vline(xintercept = pearson_original, color = "blue", linetype = "dashed") +
    geom_vline(xintercept = spearman_original, color = "red", linetype = "dashed") +
    labs(title = "Bootstrap Distributions of Correlation Coefficients",
         x = "Correlation Coefficient", y = "Density") +
    theme_minimal() +
    scale_fill_manual(values = c("Pearson" = "blue", "Spearman" = "red"))
```
## Practice 3: Estimate the value of pi
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n_points = 10000
x <- runif(n_points, -1, 1)
y <- runif(n_points, -1, 1)
inside_circle <- (x^2 + y^2) <= 1
pi_estimate <- 4 * sum(inside_circle) / n_points
cat(pi_estimate)
```
```{r, message=FALSE, echo=FALSE, warning = FALSE}
df <- data.frame(x = x, y = y, inside = inside_circle)
ggplot(df, aes(x = x, y = y, color = inside)) +
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "red", "TRUE" = "blue"),
                     labels = c("Outside", "Inside")) +
  labs(title = paste("Monte Carlo Estimation of 蟺 =", round(pi_estimate, 4)),
       x = "X", y = "Y", color = "Position") +
  theme_minimal() +
  coord_fixed()
```
## Practice 4:
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
n = 100
n_sim = 1000
theta = 2
estimates <- numeric(n_sim)
estimates_c <- numeric(n_sim)
for(i in 1:n_sim) {
  x <- rnorm(n, 0, theta)
  mixture <- runif(n) < 0.9
  x_c <- ifelse(mixture, rnorm(n, 0, 2), rnorm(n, 0, 4))
  estimates[i] <- (1/n * sum(abs(x - mean(x))))
  estimates_c[i] <- (1/n * sum(abs(x_c - mean(x_c))))
}
true_value <- theta * sqrt(2/pi)
true_value_c <- 0.9*2*sqrt(2/pi) + 0.1*4*sqrt(2/pi)
bias <- mean(estimates) - true_value
variance <- var(estimates)
mse <- mean((estimates - true_value)^2)
bias_c <- mean(estimates_c) - true_value_c
variance_c <- var(estimates_c)
mse_c <- mean((estimates_c - true_value_c)^2)
results_df <- data.frame(
  Distribution = c("Normal", "Contaminated"),
  Theory = c(true_value, true_value_c),
  Bias = c(bias, bias_c),
  Variance = c(variance, variance_c),
  MSE = c(mse, mse_c)
)
print(results_df)
```
This estimated quantity is the Mean Absolute Deviation (MAD), which is more robust to outliers compared to the standard deviation. By comparing its performance under normal distribution and contaminated distribution, we can observe its robustness when outliers are present.